Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2022-07-30 elsarticle/elsarticle.tex
change plot legend

 Due to this linearization, the infection/recovery ratio, $\phi(t)$, will have to change as a function of time to accommodate for the boundary conditions. And since $I$ is a function of $Y$, we can combine Eq. \ref{eq:linSIR} via an instantaneous relative growth rate, $\gamma(t) = dY(t) /(Ydt) = \beta I / Y$, in turn parameterized by a scaling factor, $\theta$, representing the shape of the growth,
\begin{subequations}
\label{eq:rypdal}
\begin{align}
\frac{dY}{dt} =& \gamma(t) Y(t) \label{eq:rypdalODE}\\
\gamma(t) =& \frac{\gamma_{\infty}}{\theta}\left[1 - \left(\frac{Y}{Y_{\infty}}\right)^{\theta} \right] \label{eq:rypdalGamma},
\end{align}
\end{subequations}
where $\gamma_{\infty} = \gamma(t\rightarrow \infty)$. This parametrization is the commonly used Richard's growth curve \citep{richards1959flexible}, also called $\theta$-logistic growth, and has been used by others \citep{wu2020generalized}. Although not immediately justified in the communicable disease theory, one could imagine that $\theta$ represents non-linear network behavior \citep{petroni2020logistic}. Note that at $\theta=1$, the traditional logistic growth curve is obtained, while at $\theta\rightarrow \infty$ we recover the exponential (Malthusian) explosion.

@@

\subsection{The extended SIR model family is almost Gompertzian}
Without the linearization and the $\theta$-parametrization of the Richard's model, one can still almost obtain Gompertzian growth from the communicable disease models. Under a double-log transform of Eq. \ref{eq:gomp_solution}, the growth curve is a simple straight line which can be seen as a tell-tale sign of a Gompertz curve,
\begin{equation}
\label{eq:GOMP_D}
\ln{(\ln{(Y_{\infty}/Y)})} = -\beta t + k,
\end{equation}
with $k=\ln{\ln{\frac{Y_{\infty}}{Y_{0}}}}$. To show this, we follow Carletti et al. (2020) \cite{carletti2020covid} and consider the extended version of the SIR the model by including a group of diseased, $D(t)$, such that $N = S(t) + I(t) + R(t) + D(t)$, also called the SIRD model. This requires an addition to our original set of equations \ref{eq:SIR}, with an extra equation for the diseased group's growth at some rate $\delta$ relative to the current infected group,
\begin{equation}
\frac{dD}{dt} = \delta I.
\end{equation}

@@

Thus, the emergence of the Gompertz curve at the macroscopic level suggests that the system is correlated, or coherent, presumably as a result of the simultaneous exposure to the same underlying stressor, but also due to the implied log-normal nature of the microscopic entities, where multiplication replaces addition as the aggregating operator \citep{zhang1994log}.
We can now further appreciate Richard's parametrization as a transition from non-collaborative to collaborative growth as $\theta \rightarrow 0$. 
This feature of $\theta$ was also obtained by Petroni et al. (2020) \cite{petroni2020logistic} by interpreting the $\theta-$logistic growth rate in Eq. \ref{eq:rypdal} as non-linear resource availability dependent on the overall magnitude, $Y(t)$, with growth at $\theta \rightarrow 0$ named ``maximally coherent''. 
Molski and Konarski (2003) \cite{molski2003coherent} made a similar observation that Gompertzian growth is the coherent state in a quantum mechanical system with a time-dependent potential, an interpretation which sheds further light on the temporal nature of the postulated stress response.
This quantum mechanical system has also been used to describe coherent energy states of diatomic molecules in space \citep{morse1929diatomic}. 
In the field of quantum physics, \emph{coherence} is a well-defined mathematical property first explored by Glauber (1963) \cite{glauber1963coherent} in the context of electromagnetic fields. The fact that we observe the Gompertz curve in both the microscopic quantum scales and the macroscopic national scales is noteworthy, and suggests that both systems share commonalities and means of communication.

@@

\insertPdfFig{hypergeometric}{Comparison relative growth rate in the Richard's $\theta$-logistic model (Eq. \ref{eq:rypdal}), Gompertz model (Eq. \ref{eq:medianGomp}), and the augmented causality-reversed SIR model (Eq. \ref{eq:modLogistic}), all as a function of the abundance variable. Both dependent and independent variables are normalized so that final abundance (or size) is set to unity. The Richard's $\theta$-parameter is annotated in brackets in the legend, in addition to the number of terms used in the modified SIR model.}

\section{Conclusion}
In conclusion, we have shown that Gompertzian growth follows from infinite interactions between the susceptible and infected states, and that the perceived pathogen travels at infinite speed throughout the population, rejecting the possibility of a disease propagation stage through a perceived transmission network. In this vein, Richard's parameter, $\theta$, is related although not identical to the number of higher-order interactions with the susceptible and the infected population in a SIR model \citep{richards1959flexible}, where infinite interactions corresponds to $\theta\rightarrow 0$.

>>>>

 Due to this linearization, the infection/recovery ratio, $\phi(t)$, will have to change as a function of time to accommodate for the boundary conditions. And since $I$ is a function of $Y$, we can combine Eq. \ref{eq:linSIR} via an instantaneous relative growth rate, $\gamma(t) = dY(t) /(Ydt) = \beta I / Y$, in turn parameterized by a scaling factor, $\theta$, representing the shape of the growth,
\begin{subequations}
\label{eq:rypdal}
\begin{align}
\frac{dY}{dt} =& \gamma(t) Y(t) \label{eq:rypdalODE}\\
\gamma(t) =& \frac{\gamma_{\infty}}{\theta}\left[1 - \left(\frac{Y}{Y_{\infty}}\right)^{\theta} \right] \label{eq:rypdalGamma},
\end{align}
\end{subequations}
where $\gamma_{\infty} = \gamma(t\rightarrow \infty)$. This parametrization is the commonly used Richards growth curve \citep{richards1959flexible}, also called $\theta$-logistic growth, and has been used by others \citep{wu2020generalized}. Although not immediately justified in the communicable disease theory, one could imagine that $\theta$ represents non-linear network behavior \citep{petroni2020logistic}. Note that at $\theta=1$, the traditional logistic growth curve is obtained, while at $\theta\rightarrow \infty$ we recover the exponential (Malthusian) explosion.

@@

\subsection{The extended SIR model family is almost Gompertzian}
Without the linearization and the $\theta$-parametrization of the Richards model, one can still almost obtain Gompertzian growth from the communicable disease models. Under a double-log transform of Eq. \ref{eq:gomp_solution}, the growth curve is a simple straight line which can be seen as a tell-tale sign of a Gompertz curve,
\begin{equation}
\label{eq:GOMP_D}
\ln{(\ln{(Y_{\infty}/Y)})} = -\beta t + k,
\end{equation}
with $k=\ln{\ln{\frac{Y_{\infty}}{Y_{0}}}}$. To show this, we follow Carletti et al. (2020) \cite{carletti2020covid} and consider the extended version of the SIR the model by including a group of diseased, $D(t)$, such that $N = S(t) + I(t) + R(t) + D(t)$, also called the SIRD model. This requires an addition to our original set of equations \ref{eq:SIR}, with an extra equation for the diseased group's growth at some rate $\delta$ relative to the current infected group,
\begin{equation}
\frac{dD}{dt} = \delta I.
\end{equation}

@@

Thus, the emergence of the Gompertz curve at the macroscopic level suggests that the system is correlated, or coherent, presumably as a result of the simultaneous exposure to the same underlying stressor, but also due to the implied log-normal nature of the microscopic entities, where multiplication replaces addition as the aggregating operator \citep{zhang1994log}.
We can now further appreciate the Richards parametrization as a transition from non-collaborative to collaborative growth as $\theta \rightarrow 0$. 
This feature of $\theta$ was also obtained by Petroni et al. (2020) \cite{petroni2020logistic} by interpreting the $\theta-$logistic growth rate in Eq. \ref{eq:rypdal} as non-linear resource availability dependent on the overall magnitude, $Y(t)$, with growth at $\theta \rightarrow 0$ named ``maximally coherent''. 
Molski and Konarski (2003) \cite{molski2003coherent} made a similar observation that Gompertzian growth is the coherent state in a quantum mechanical system with a time-dependent potential, an interpretation which sheds further light on the temporal nature of the postulated stress response.
This quantum mechanical system has also been used to describe coherent energy states of diatomic molecules in space \citep{morse1929diatomic}. 
In the field of quantum physics, \emph{coherence} is a well-defined mathematical property first explored by Glauber (1963) \cite{glauber1963coherent} in the context of electromagnetic fields. The fact that we observe the Gompertz curve in both the microscopic quantum scales and the macroscopic national scales is noteworthy, and suggests that both systems share commonalities and means of communication.

@@

\insertPdfFig{hypergeometric}{Comparison relative growth rate in the Richards $\theta$-logistic model (Eq. \ref{eq:rypdal}), Gompertz model (Eq. \ref{eq:medianGomp}), and the augmented causality-reversed SIR model (Eq. \ref{eq:modLogistic}), all as a function of the abundance variable. Both dependent and independent variables are normalized so that final abundance (or size) is set to unity. The Richards $\theta$-parameter is annotated in brackets in the legend, in addition to the number of terms used in the modified SIR model.}

\section{Conclusion}
In conclusion, we have shown that Gompertzian growth follows from infinite interactions between the susceptible and infected states, and that the perceived pathogen travels at infinite speed throughout the population, rejecting the possibility of a disease propagation stage through a perceived transmission network. In this vein, the Richards parameter, $\theta$, is related although not identical to the number of higher-order interactions with the susceptible and the infected population in a SIR model \citep{richards1959flexible}, where infinite interactions corresponds to $\theta\rightarrow 0$.
